Dispersion distance and the matter distribution of the Universe in dispersion space

### 2015-06-04 writeup/dm_space_distortions.tex
Added mention of SKA.

    We propose that `standard pings', brief broadband radio impulses, can be
    used to study the three-dimensional clustering of matter in the Universe
    even in the absence of redshift information.  The dispersion of radio waves
    as they travel through the intervening plasma can,
    like redshift, be used as a
    cosmological distance measure.  Due to inhomogeneities in the electron
    density along the line-of-sight, dispersion is an imperfect proxy for
    radial distance and we show that this leads to calculable dispersion-space
    distortions in the apparent clustering of sources. Fast radio bursts (FRBs)
    are a new class of radio transients that are the prototypical standard
    ping, and due to their high observed dispersion,
    have been interpreted as originating at cosmological distances.  The rate
    of fast radio bursts has been estimated to be several thousand over the
    whole sky per day, and if cosmological, the sources of these events should
    trace the large-scale structure of the Universe.  We calculate the
    dispersion-space power spectra for a simple model where electrons and FRBs
    are biased tracers of the large-scale structure of the Universe and show
    that the clustering signal could be measured using as few as
    $10000$ events. Such a survey is in-line with what may be achieved with
    upcoming wide-field radio telescopes such as the Canadian
    Hydrogen Intensity Mapping Experiment (CHIME).
\end{abstract}

@@

We have
shown in Figure~\ref{f:sensitivity}
that a survey detecting $10000$ fast radio bursts of cosmological
origin could detect the
clustering signal. This is in rough
approximation to what might be achieved by the Canadian Hydrogen Intensity
Mapping Experiment (CHIME)\footnote{\url{http://chime.phas.ubc.ca/}}
or other upcoming wide-field telescopes.
Looking forward, dispersion-space
clustering could become a precision probe of large-scale structure should
observational factors turn out to be favourable compared to spectroscopic
surveys. The baryon acoustic oscillation feature can be clearly seen in
Figure~\ref{f:terms} in the cross term and both
local-term curves. It may also be possible to extract the BAO feature
from the integral
term using tomography.
This however requires that the contribution to dispersion from
the source's environment and host halo be small compared to the
contribution from the IGM.

>>>>

    We propose that `standard pings', brief broadband radio impulses, can be
    used to study the three-dimensional clustering of matter in the Universe
    even in the absence of redshift information.  The dispersion of radio waves
    as they travel through the intervening plasma can,
    like redshift, be used as a
    cosmological distance measure.  Due to inhomogeneities in the electron
    density along the line-of-sight, dispersion is an imperfect proxy for
    radial distance and we show that this leads to calculable dispersion-space
    distortions in the apparent clustering of sources. Fast radio bursts (FRBs)
    are a new class of radio transients that are the prototypical standard
    ping, and due to their high observed dispersion,
    have been interpreted as originating at cosmological distances.  The rate
    of fast radio bursts has been estimated to be several thousand over the
    whole sky per day, and if cosmological, the sources of these events should
    trace the large-scale structure of the Universe.  We calculate the
    dispersion-space power spectra for a simple model where electrons and FRBs
    are biased tracers of the large-scale structure of the Universe and show
    that the clustering signal could be measured using as few as
    $10000$ events. Such a survey is in-line with what may be achieved with
    upcoming wide-field radio telescopes.
\end{abstract}

@@

We have
shown in Figure~\ref{f:sensitivity}
that a survey detecting $10000$ fast radio bursts of cosmological
origin could detect the
clustering signal. This is in rough
approximation to what might be achieved by the Canadian Hydrogen Intensity
Mapping Experiment (CHIME)\footnote{\url{http://chime.phas.ubc.ca/}}, the
Square Kilometer Array\footnote{\url{http://www.skatelescope.org/}},
or other upcoming wide-field telescopes.
Looking forward, dispersion-space
clustering could become a precision probe of large-scale structure should
observational factors turn out to be favourable compared to spectroscopic
surveys. The baryon acoustic oscillation feature can be clearly seen in
Figure~\ref{f:terms} in the cross term and both
local-term curves. It may also be possible to extract the BAO feature
from the integral
term using tomography.
This however requires that the contribution to dispersion from
the source's environment and host halo be small compared to the
contribution from the IGM.
